---
title: Superbowl Ads
author: by Dorian's Gs
output: github_document
---

```{r load-packages, echo = FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(fs)
library(ggplot2)
library(dplyr)
library(ggridges)
library(forcats)
library(colorspace)
library(RColorBrewer)
```

```{r load-data, echo = FALSE}
superbowl<- read.csv("data/superbowl.csv")
```

## Introduction

The dataset `superbowl` is created by FiveThirtyEight, originally collected through superbowl-ads.com. There are 247 different superbowl commercials with 25 different variables, but observe only 10 different ads Additionally, the dataset includes defining characteristics of superbowl ads from popular brands like Toyota and Bud Light. 

There are different types of variables included in `superbowl`: 7 of them are integers, including the view_count, like_count, dislike_count. 11 variables are characters, including brand, title, and description. 7 variables are logical, which shows either TRUE or FALSE for some categories, including whether or not the ad is funny, patriotic, includes animals, and includes sex. These logical variables were determined by the FiveThirtyEight team as they watched and categorized all of the advertisements.

Using the dataset `superbowl`, we answer the following questions: 

## 1. What factors contribute to the most viewed ads and has the relationship between those factors and the views changed over time?

### Introduction

We investigated how including different factors such as animals, celebrity, and sexuality affect the number of views on an ad. Even though the dataset provides a lot more variables related to ad characteristics, we chose to focus on these three because we think they have the most potential to influence view counts based on our personal experiences and observations. For the second part of the question, we will primarily investigate how the trends and relationships we explored in the first part have changed over the years.

We were interested in this question because we’ve observed how influential ads can be if they include characteristics that appeal to our emotions, such as our adoration for a cute puppy, longing to be like a popstar, or simply our sexual appetite. However, we are unsure which emotional appeal would foster more views and think it is important for advertisement agencies to know. Additionally, we observed that certain ad appeals, especially use of sexuality, have grown in popularity over the years, and would like to observe such trends with our data analysis.

The relevant variables include:

* `animals`: logical variable, whether or not the ad includes an animal

* `celebrity`: logical variable, whether or not the ad includes a celebrity

* `use_sex`: logical variable, whether or not the ad includes sexuality

* `view_count`: numerical variable, number of views on the ad

* `year`: numerical variable, year the ad was released


### Approach

(1-2 paragraphs)
Describe what types of plots you are going to make to address your question.
For each plot, provide a clear explanation as to why this plot (e.g. boxplot, barplot, histogram, etc.) is best for providing the information you are asking about. 
The two plots should be of different types, and at least one of the two plots needs to use either color mapping or facets.

#### Plot 1: 


#### Plot 2:
For the second plot, we wanted to look more specifically  at the relationship between view count and the number of interactions a video gets. We defined interactions as the number of likes, dislikes and comments a video got. We wanted to plot both `view_count` and `interactions` as continuous variables.  A unique and challenging aspect of our dataset was the range of view count was so large that it was nearly impossible to  visualize any interaction with just one plot. To solve that problem, we decided to facet our plot by the 6 categories for view count  that we created for part 1. We then freed the scales of each axis to make it much easier to visualize the trend between interactions and view count. Since we are primarily interested in the ratio of the views to interactions , the difference in scale between the x-axis(the view count) and the y-axis(number of interactions)  is actually useful for getting a sense of the difference in factors between the two variables


In order to find the best plot, we experimented with  scatter plots, line plots, step plots, and area plots to see which plot  was the most informative. There were too little variables for the scatter plot to be helpful. It was also hard to follow along with the variable. We found that a line plot or step  by itself was to fragmented and the connection between points zig-zagged a lot and made it hard to also see the overall trend. The combined plot shows each points but the area also makes it much easier to trace the trend for each view category. We then added labels, captions and made a few styling choices to get to our final presentation. 





### Analysis

(2-3 code blocks, 2 figures, text/code comments as needed)
In this section, provide the code that generates your plots. Use scale functions to provide nice axis labels and guides.
You are welcome to use theme functions to customize the appearance of your plot, but you are not required to do so.
All plots must be made with ggplot2. Do not use base R or lattice plotting functions.



### Discussion

(1-3 paragraphs)
In the Discussion section, interpret the results of your analysis. Identify any trends revealed (or not revealed) by the plots.
Speculate about why the data looks the way it does.

## 2. What is the relationship between popularity of a Superbowl Ad and how well it is interacted with?

### Introduction

We were interested in examining the relationship between different popularity and interaction metrics of the superbowl ads because we want to more clearly understand what a “successful ad” entails. Some agencies could prefer more comments and likes to an ad to show that consumers are engaging with the content instead of merely looking at view count. If so, that agency would create a commercial that fosters more likes than pure views. 

The relevant variables are all numerical and include: `view_count`, `like_count`, `dislike_count`, `comment_count`, and `favorite count`. The `view_count` is the measure for popularity, while the other variables define interaction with the ad. 


### Approach

For the first plot, we examined the relationship between `view_count` and the like percentage of the video using a density ridge plot. More specifically, we wanted to observe whether or not the amount of views on a Superbowl ad correlates with a higher or lower like percentage. In order to plot the like percentage, we created a new variable `ratio` which is the `like_count` divided by the sum of `like_count` and `dislike_count`, and plotted the x-axis labels to show in percentage (since the like ratio on Youtube is shown in percentage form). We categorized `view_count` into 6 distinct categories (ranging from few to viral). We chose a density ridge plot because it clearly shows changes in like percentage while comparing these changes by the view count category. The peaks of the denstiy ridges make comparisons of like percentages among view categories simple to track. Additionally, since we categorized view count into 6 distinct groups, a categorical y-axis was necessary.



(1-2 paragraphs)
Describe what types of plots you are going to make to address your question.
For each plot, provide a clear explanation as to why this plot (e.g. boxplot, barplot, histogram, etc.) is best for providing the information you are asking about. 
The two plots should be of different types, and at least one of the two plots needs to use either color mapping or facets.

### Analysis

(2-3 code blocks, 2 figures, text/code comments as needed)
In this section, provide the code that generates your plots. Use scale functions to provide nice axis labels and guides.
You are welcome to use theme functions to customize the appearance of your plot, but you are not required to do so.
All plots must be made with ggplot2. Do not use base R or lattice plotting functions.

### Discussion

(1-3 paragraphs)
In the Discussion section, interpret the results of your analysis. Identify any trends revealed (or not revealed) by the plots.
Speculate about why the data looks the way it does.


#### Question 1



#### Question 2. 


** Plot 2.**

The area plot of view count by number of interactions reveals quite a few interesting results. The first is that on average, the number of views for ads is larger by a factor of about 100 that the number of interactions on that ad. On average, about 100 views resulted in 1 interaction with the video. 
Our guess prior to seeing the results of our visualization was that as the number of views increased, the number of proportions would also increase on a statistically significant rate. In terms of our data that means that we expected each higher category of views would have a significantly high level of interaction. As an example, we expected that proportionally, the number of view count to interactions would be higher for ads with “Viral Views” compared to “Mega Views” and ads with “Mega Views” would be higher that graphs with “Many Views”. Our results were turned out to be a bit more complicated than that. We found that only ads on either extreme end followed this trend. So ads with a “few views” had much  a smaller views to interaction ratio than videos in all other categories and ads with “viral views” had much bigger views to interaction ratio than videos in all other “categories.” For videos in the middle categories("Some Views","Moderate Views","Many Views" and "High Views"), proportion of view counts to interactions seemed to be much more random and didn’t follow any discernible patterns. Each panel also had no discernible trend in the way they were plotted unlike the “few” and “viral” categories which had a discernible linear pattern each. 


## Presentation

Our presentation can be found [here](https://vizdata-f21.github.io/project-1-dorian_s_gs/presentation/presentation.html#1).

## Data 
Best R and Mehta D, 2021, *Super Bowl Ads*, Electronic Dataset, Github Repository, Retrieved September 13, 2021 <https://github.com/fivethirtyeight/superbowl-ads>. 

## References

1. Data Source: Our data source is from [The Tidy Tuesday Project](https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-03-02/readme.md) by way of [FiveThirtyEight's Github Repository](https://github.com/fivethirtyeight/superbowl-ads).
2. Viral Reference: https://www.forbes.com/sites/robertwynne/2018/03/09/there-are-no-guarantees-or-exact-statistics-for-going-viral/?sh=6224f76a5e8c

Presentation Images:

3. Title image: https://www.pinterest.com/pin/746542075709663730/
4. Youtube logo: Youtube
5. Tv image: FiveThirtyEight
6. Logos directly from: Budweiser, Bud Light, Pepsi, Doritos, Coca-Cola, Hyundai, E Trade, Kia, Toyota, and NFL
7. Youtube interaction bar: https://www.tubefilter.com/2021/03/30/youtube-experiment-hiding-public-dislikes-creator-well-being/
